caDDS: Community Assisted Distributed Database for Sequences

### 2020-06-26 Paper2020ACM/MeetingJune25.tex
Revision June 26 3pm

Whole Paper:
1) Make it 2 column format
2) Add sir El King Morado - UP Philippine Genome Center
3) Remove references to page numbers
4) Put figures on top of a column (see what sir el king sent)

Abstract
- okay content
- fix typoL Data nods

@@

-Intro Pg 2
- 3,000,000,000 -> 3 billion
- Shorten HGP section. Combine HGP + NGS to one paragraph.
- Fix typo: Genomics is is

@@

Acknowledgements -> Remove from Appendix

Biblio
- Make it APA format

>>>>

Whole Paper:
check 1) Make it 2 column format
check 2) Add sir El King Morado - UP Philippine Genome Center
check 3) Remove references to page numbers
check 4) Put figures on top of a column (see what sir el king sent)

Abstract
- okay content
check - fix typoL Data nods

@@

-Intro Pg 2
check - 3,000,000,000 -> 3 billion
- Shorten HGP section. Combine HGP + NGS to one paragraph.
check - Fix typo: Genomics is is

@@

check Acknowledgements -> Remove from Appendix

Biblio
check - Make it APA format

### 2020-06-26 Paper2020ACM/Paper2020.tex
Thesis submission June 25 6pm
Revision June 26 3pm

%%
%% This is file `sample-acmsmall.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[acmsmall]{acmart}

@@

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

@@

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Carl Araya}
\authornote{Both authors contributed equally to this research.}
\email{jjazcarraga@up.edu.ph}
\affiliation{%
  \institution{University of the Philippines - Diliman}
  \city{Quezon}
  \country{Philippines}
}

\author{Jose Carlos Rodrigo J. Azcarraga}
\authornotemark[1]
\affiliation{%
  \institution{University of the Philippines - Diliman}
  \city{Quezon}
  \country{Philippines}
}
\email{jjazcarraga@up.edu.ph}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Genomic data the past few years have grown exponentially with the technology barely keeping up. The problem faced by genome researchers is the large data set and difficulty transferring the files. Previous distributed databases are either not meant for genome data, or difficult to replicate. This paper lays the groundwork for a distributed database that is designed to easily scale to accommodate bigger storage, and also reduces the data speed bottleneck faced by centralized databases. The proposed system has a master node and data nodes. Master node handles administrative work while the Data nods contain all sequence data. Data nodes should theoretically increase data transfer speeds. While current databases(e.g. NCBI) have more data, more users, and a live platform. The proposed system is aimed towards a smaller community, with more frequent and localized data, which the data storage can easily be expanded as the need arises. The proposed system is also theoretically faster as long as there are multiple working data nodes. A recommendation is to implement the system and make the code public for the use of other communities.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
The goal of this introduction is to explain the idea of genomes, their size and magnitude. Inside every human body are the instructions to create and maintain it. These instructions are found inside the genome. The instructions inside the human genome are incredibly long. For example, entire human genome, inside a singular cell, when laid out is around 2 meters in length.\cite{ency_sci_tech}

@@

\subsection{Definition of Genomics}
Genomics is is the study of wholes sets of genes and their interactions\cite[~p.437]{campbell}. The efforts to study this field has led to volumes of data that need to be analyzed. Which has also spurred the field bioinformatics, which aims to use computers to aid the analysis and storage of biological data. Institutions such as the National Institute of Health and National Library of Medicine, have led this field and built a website to encourage collaboration online through their online portal called NCBI (National Center for Biotechnology Information). \cite[~p.438]{campbell} This website will be discussed later in more detail in Section \ref{NCBI}.

@@

Another way to see the way genomic data is large is by looking at the \textit{Human Genome Project (HGP)}. The project's aim was to sequence the entire human genome. At the time it needed a lot of collaborations, a lot of financing, and new computational tools (piecing everything together). Back in 2003 it was a milestone because the first human genome \cite{introgenomics} had been first completely read. It took the project ten years of work, a dozen institutions, and 3,000,000,000 dollars to finish. The technology for sequencing and genomics was still at its infancy.

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.65\textwidth]{images/human-gen-cost.jpg} 
\caption{Cost per Genome Data Over Time}
\label{fig:human_gen_cost_fig}
\end{figure}

To view these technology developments over time, we look at the Figure~\ref{fig:human_gen_cost_fig} on Page~\pageref{fig:human_gen_cost_fig} \cite{genomics-cost}, we note here the expensive cost at the initial stages of sequencing the human genome. But as time progresses, the cost substantially went down to what it is now. The big fall of the prices around 2004 - 2005 corresponds to the introduction of the Next Generation Sequencers.

@@

\begin{figure}[h]
\caption{Megabase Cost Over Time}
\centering
\includegraphics[width=0.65\textwidth]{images/seq-cost.jpeg} 
\label{fig:megabase_cost_fig}
\end{figure}

Similar to Figure~\ref{fig:human_gen_cost_fig} on Page~\pageref{fig:human_gen_cost_fig}, we also note the same pattern for the cost of sequencing (not just humans) in Figure~\ref{fig:megabase_cost_fig}\cite{genomics-cost}. After the human genome project, many more whole genome projects have been completed. And the cost of sequencing dropped 1000 fold from 2008 to 2012 alone \cite{bon_compression}. This lowering of cost, again corresponds to the idea of sequencing becoming more available. Increasing the number of research projects, therefore contributing to the growing number again of genomic data.

@@

% CID PROGRESS ENDS HEREE FOR MAY 29 2020

@@

% CID PROGRESS STARTS HEREE FOR JUNE 2 2020

\begin{table}[h]
    \caption{Comparisons of Different Databases and Technologies}
    \label{table:database_comparison_table}
    \begin{tabular}{ccccc}
    \toprule
    System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ 
    \midrule
    BioTorrents & P2P           & No         & Depends                    & Biological  \\ 
    SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ 
    PeerDB      & P2P           & No         & Depends                    & General      \\ 
    NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\
    Dat Protocol    & P2P      & Yes        & Depends           & General  \\ 
    \bottomrule
    \end{tabular}
\end{table}

@@

Table~\ref{table:database_comparison_table} on page ~\pageref{table:database_comparison_table} compares the different technologies studied. An important point to note is that a majority of the databases are not available. This will also be used later on to describe what are the main problems of the different databases.

@@

% CID PROGRESS ENDS HEREE FOR JUNE 2 2020

% CID PROGRESS STARTS HEREE FOR JUNE 3 2020
% REDID FORMAT
% CID PROGRESS ENDS HEREE FOR JUNE 3 2020

% CID PROGRESS STARTS HEREE FOR JUNE 4 2020

@@

% CID PROGRESS ENDS HEREE FOR JUNE 4 2020

\begin{figure}[h]
\caption{App network diagram}
\centering
\includegraphics[width=0.35\textwidth]{images/thesis1.png} 
\end{figure}

\begin{figure}[h]
\caption{POST: Uploading a sequence to the master node}
\centering
\includegraphics[width=0.35\textwidth]{images/thesis3.png} 
\end{figure}

\begin{figure}[h]
\caption{GET: Downloading a sequence}
\centering
\includegraphics[width=0.35\textwidth]{images/thesis4.png} 
\end{figure}

%CID: Explain the data model and data permissions

\begin{table}[h]
\caption{Data Model to be Used in the System}
\label{table:data_model_table}
\begin{tabular}{cl}
    \toprule
    key & description \\
    \midrule
    seq\textunderscore id & unique ID assigned to all sequences \\
    organism & organism where the sequence was extracted from \\
    quality & the rating of the data based on the number of downloads or another criteria \\
    uploader & user that uploaded the data \\
    institute & institute where the uploader or sequence was taken from \\
    upload\textunderscore date & date the data was uploaded to the database \\
    last\textunderscore modified & the date the metadata or data was edited \\
    data\textunderscore nodes & list of data nodes that this data is contained to \\
    \bottomrule
\end{tabular}
\end{table}

Table~\ref{table:data_model_table} on Page~\pageref{table:data_model_table} aims to show the metadata that will be initially stored in the database.

%CID: Data Permissions

\begin{table}[h]
\caption{Data Permissions and Users in the System}
\label{table:data_perm_table}
\begin{tabular}{cl}
    \toprule
    user & description \\
    \midrule
   Moderator & Has ability to upload, download, verify \\
    Registered User & Has ability to upload, download \\
    Guest & Has ability to download \\
   \bottomrule 
\end{tabular}
\end{table}

Table~\ref{table:data_perm_table} on Page~\pageref{table:data_perm_table} shows the different set of users and the permissions allowed to them in the system. The moderator is a special user who has the ability to not allow some data to be uploaded. This is to make sure there is a means to verifying the data being uploaded. All registered users can upload and download. All other users, which are called guests, have an ability to download data.

@@

%\begin{figure}[h]
%  \centering
%  \includegraphics[width=\linewidth]{sample-franklin}
%  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
%    Ewing, Inc. [Public domain], via Wikimedia
%    Commons. (\url{https://goo.gl/VLCRBB}).}
%  \Description{A woman and a girl in white dresses sit in an open car.}
%\end{figure}

\section{Appendices}

@@

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{bibfile}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

>>>>

%%
%% This is file `sample-acmsmall.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[acmsmall]{acmart}
%\documentclass[manuscript,screen]{acmart}
\usepackage{multicol, caption}
\usepackage{placeins}

\newenvironment{Figure}
  {\par\medskip\noindent\minipage{\linewidth}}
  {\endminipage\par\medskip}

@@

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%\citestyle{acmauthoryear}

@@

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{El King Morado}
\authornote{Adviser}
\email{elkingmoradoking@gmail.com}
\affiliation{%
  \institution{University of the Philippines - Philippine Genome Center}
  \city{Quezon}
  \country{Philippines}
}

\author{Carl Araya}
\email{carl.araya256@gmail.com}
\affiliation{%
  \institution{University of the Philippines - Diliman}
  \city{Quezon}
  \country{Philippines}
}

\author{Jose Carlos Rodrigo J. Azcarraga}
\email{jjazcarraga@up.edu.ph}
\affiliation{%
  \institution{University of the Philippines - Diliman}
  \city{Quezon}
  \country{Philippines}
}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Genomic data the past few years have grown exponentially with the technology barely keeping up. The problem faced by genome researchers is the large data set and difficulty transferring the files. Previous distributed databases are either not meant for genome data, or difficult to replicate. This paper lays the groundwork for a distributed database that is designed to easily scale to accommodate bigger storage, and also reduces the data speed bottleneck faced by centralized databases. The proposed system has a master node and data nodes. Master node handles administrative work while the Data nodes contain all sequence data. Data nodes should theoretically increase data transfer speeds. While current databases(e.g. NCBI) have more data, more users, and a live platform. The proposed system is aimed towards a smaller community, with more frequent and localized data, which the data storage can easily be expanded as the need arises. The proposed system is also theoretically faster as long as there are multiple working data nodes. A recommendation is to implement the system and make the code public for the use of other communities.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{multicols}{2}

\section{Introduction}
Inside every human body are the instructions to create and maintain it. These instructions are found inside the genome. The instructions inside the human genome are incredibly long. For example, entire human genome, inside a singular cell, when laid out is around 2 meters in length.\cite{ency_sci_tech}

@@

\subsection{Definition of Genomics}
Genomics is the study of whole sets of genes and their interactions\cite[~p.437]{campbell}. The efforts to study this field has led to volumes of data that need to be analyzed. Which has also spurred the field bioinformatics, which aims to use computers to aid the analysis and storage of biological data. Institutions such as the National Institute of Health and National Library of Medicine, have led this field and built a website to encourage collaboration online through their online portal called NCBI (National Center for Biotechnology Information). \cite[~p.438]{campbell} This website will be discussed later in more detail in Section \ref{NCBI}.

@@

Another way to see the way genomic data is large is by looking at the \textit{Human Genome Project (HGP)}. The project's aim was to sequence the entire human genome. At the time it needed a lot of collaborations, a lot of financing, and new computational tools (piecing everything together). Back in 2003 it was a milestone because the first human genome \cite{introgenomics} had been first completely read. It took the project ten years of work, a dozen institutions, and 3 billion dollars to finish. The technology for sequencing and genomics was still at its infancy.

@@

\begin{Figure}
\centering
\includegraphics[width=0.80\linewidth]{images/human-gen-cost.jpg} 
\captionof{figure}{Cost per Genome Data Over Time}
\label{fig:human_gen_cost_fig}
\end{Figure}

To view these technology developments over time, we look at the Figure~\ref{fig:human_gen_cost_fig} \cite{genomics-cost}, we note here the expensive cost at the initial stages of sequencing the human genome. But as time progresses, the cost substantially went down to what it is now. The big fall of the prices around 2004 - 2005 corresponds to the introduction of the Next Generation Sequencers.

@@

\begin{Figure}
\captionof{figure}{Megabase Cost Over Time}
\centering
\includegraphics[width=0.80\linewidth]{images/seq-cost.jpeg} 
\label{fig:megabase_cost_fig}
\end{Figure}

Similar to Figure~\ref{fig:human_gen_cost_fig}, we also note the same pattern for the cost of sequencing (not just humans) in Figure~\ref{fig:megabase_cost_fig}\cite{genomics-cost}. After the human genome project, many more whole genome projects have been completed. And the cost of sequencing dropped 1000 fold from 2008 to 2012 alone \cite{bon_compression}. This lowering of cost, again corresponds to the idea of sequencing becoming more available. Increasing the number of research projects, therefore contributing to the growing number again of genomic data.

@@

@@

\begin{table*}[ht]
    \caption{Comparisons of Different Databases and Technologies}
    \label{table:database_comparison_table}
    \begin{tabular}{ccccc}
    \toprule
    System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ 
    \midrule
    BioTorrents & P2P           & No         & Depends                    & Biological  \\ 
    SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ 
    PeerDB      & P2P           & No         & Depends                    & General      \\ 
    NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\
    Dat Protocol    & P2P      & Yes        & Depends           & General  \\ 
    \bottomrule
    \end{tabular}
\end{table*}

@@

Table~\ref{table:database_comparison_table} compares the different technologies studied. An important point to note is that a majority of the databases are not available. This will also be used later on to describe what are the main problems of the different databases.

@@

@@

\begin{Figure}
\captionof{figure}{App network diagram}
\centering
\includegraphics[width=0.8\linewidth]{images/thesis1.png} 
\end{Figure}

\begin{Figure}
\captionof{figure}{POST: Uploading a sequence}
\centering
\includegraphics[width=0.80\linewidth]{images/thesis1-Page-3.png} 
\end{Figure}

\begin{Figure}
\captionof{figure}{GET: Downloading a sequence}
\centering
\includegraphics[width=0.80\linewidth]{images/thesis4.png} 
\end{Figure}

\begin{Figure}
\captionof{figure}{PUT: Updating a sequence}
\centering
\includegraphics[width=0.80\linewidth]{images/thesis1-Page-5.png} 
\end{Figure}

\begin{Figure}
\captionof{figure}{DELETE: Deleting a sequence}
\centering
\includegraphics[width=0.80\linewidth]{images/thesis1-zzzzz.png} 
\end{Figure}

\begin{table*}[ht]
\caption{Data Model to be Used in the System}
\label{table:data_model_table}
\begin{tabular}{cl}
    \toprule
    key & description \\
    \midrule
    seq\textunderscore id & unique ID assigned to all sequences \\
    organism & organism where the sequence was extracted from \\
    quality & the rating of the data based on the number of downloads or another criteria \\
    uploader & user that uploaded the data \\
    institute & institute where the uploader or sequence was taken from \\
    upload\textunderscore date & date the data was uploaded to the database \\
    last\textunderscore modified & the date the metadata or data was edited \\
    data\textunderscore nodes & list of data nodes that this data is contained to \\
    \bottomrule
\end{tabular}
\end{table*}

Table~\ref{table:data_model_table} aims to show the metadata that will be initially stored in the database.

%CID: Data Permissions
\begin{table*}[ht]
\caption{Data Permissions and Users in the System}
\label{table:data_perm_table}
\begin{tabular}{cl}
    \toprule
    user & description \\
    \midrule
   Moderator & Has ability to upload, download, verify \\
    Registered User & Has ability to upload, download \\
    Guest & Has ability to download \\
   \bottomrule 
\end{tabular}
\end{table*}

Table~\ref{table:data_perm_table} shows the different set of users and the permissions allowed to them in the system. The moderator is a special user who has the ability to not allow some data to be uploaded. This is to make sure there is a means to verifying the data being uploaded. All registered users can upload and download. All other users, which are called guests, have an ability to download data.

@@

@@

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
%\bibliographystyle{ACM-Reference-Format}
\bibliographystyle{newapa}
\bibliography{bibfile}

\end{multicols}{2}
